Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-03-25 demystify.tex
updates

A related reason is that you did not show enough interest in the program.
For example, your SOP did not mention any faculty or research area at the program, or you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges).

\paragraph{Red flags in Your Background} Plagiarism, cheating, or other academic dishonesty are \red{red} flags. Another one is if you have a history of jumping from one program to another (e.g., you have been in multiple different programs in short time). If you think you have red flags, you should address them in your SOP.
In general, these things are rare, but they do happen and cause concern to the adcom.

@@

In short, you \emph{do not need} to formally taking CS courses, you just need to show that you have these essential knowledge, e.g., through the mentioned ways. Many universities are well-aware that incoming graduate students might not have all the technical background, so they often have a \emph{``bridge''} courses to help students catch up.  For example, GMU has four bridge courses (Data Structures and Algorithms, Computer Systems, Discrete Math, and Programming Foundation) that incoming students can take to catch up on their CS knowledge.

>>>>

A related reason is that you did not show enough interests in the program.
For example, your SOP did not mention any faculty or research area at the program, or you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges).

\paragraph{Red flags in Your Background} Plagiarism, cheating, or other academic dishonesty are \red{red} flags. Another one is if you have a history of jumping from one program to another (e.g., you have been in multiple different programs in short time). If you think you have such issues, you should try to address them in your SOP.
In general, these things are rare, but they do happen and cause concern to the adcom.

@@

\paragraph{Core CS areas} The core CS areas that you should have knowledge in are:
\begin{itemize}
  \item Programming Foundation (e.g., programming in C/C++ or Java, OOP)
  \item Discrete Math (e.g., logic, set theory, proof techniques)
  \item Data Structures and Algorithms (e.g., linked lists, trees, sorting, searching)
  \item Computer OS or Systems (e.g., memory management, file systems, processes)
\end{itemize}

In short, you \emph{do not need} to formally taking CS courses, you just need to show that you have these essential knowledge, e.g., through the mentioned ways. Many universities are well-aware that incoming graduate students might not have all the technical background, so they often have a \emph{``bridge''} courses to help students catch up.  For example, GMU has four bridge courses corresponding to the four core areas above that incoming students can take to catch up on their CS knowledge.
